---
title: "Extracting LANDSAT max. monthly composites"
author: "Paulo N Bernardino"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  knitrBootstrap::bootstrap_document:
    theme: "simplex"
    highlight: Tomorrow Night Bright
    menu: FALSE
    theme.chooser: TRUE
    highlight.chooser: TRUE
---

# Extracting LANDSAT max. monthly composites
  
Paulo N Bernardino (paulo.nbernardino@gmail.com)

[Division Forest, Nature and Landscape, KU Leuven](https://ees.kuleuven.be/fnl/staff/index.html?intranet=u0115271)

[Laboratory of Geo-information Science and Remote Sensing, Wageningen UR](https://www.wur.nl/en/Persons/Paulo-P-Paulo-Negri-Bernardino-MSc.htm)


## Getting started

Set your working directory to the folder with the Landsat raster files. Next, the code below will install/load the required packages. 

```{r, echo=TRUE, message=FALSE, eval=TRUE}
# pkgTest is a helper function to load packages and install packages only when they are not installed yet.
pkgTest <- function(x)
{
  if (x %in% rownames(installed.packages()) == FALSE) {
    install.packages(x, dependencies= TRUE)
  }
  library(x, character.only = TRUE)
}
neededPackages <- c("zoo", "gimms", "raster", "rgdal")
for (package in neededPackages){pkgTest(package)}
```

Load the files into R.

```{r, echo=TRUE, message=FALSE}
files<-list.files(pattern="*.tif")
path<-list.files(pattern="*.tif", full.names = TRUE)
lc8<-stack(path)

## Setting masked values as NA instead of zero
lc8[lc8<=0] <- NA
```

## Creating the monthly composites

Making monthly composites, using the max per-pixel value among the available months.

```{r, echo=TRUE, message=FALSE}
## Finding corresponding months inside a year
years <- as.numeric(substr(names(lc8), star=13, stop=16))
pos_yrs<-unique(years)

## Going through all the years
out <- raster(extent(lc8[[1]]), ncols=ncol(lc8[[1]]), nrow=nrow(lc8[[1]]),
               crs=proj4string(lc8))
for (i in 1:length(pos_yrs)){
  y <- lc8[[which(years==pos_yrs[i])]] # select only data for a single year
  months <- as.numeric(substr(names(y), start=17, stop=18)) # object with 
                                                            # months' indices
  mc <- monthlyComposite(y, indices=months, fun=max) # make monthly composites
                                                     # using max value of a                                                        pixel
  
  names(mc) <- paste0(pos_yrs[i],"_",unique(months))
  out <- addLayer(out, mc)
}
```

Visulalizing some of the maps before and after compositing. Note that here I used only data between Jan-2014 and Dec-2015.

```{r, echo=TRUE, message=FALSE}
par(mfrow=c(2,2))

plot(lc8[[6]], main="May 2014 (before)")
plot(out[[4]], main="May 2014 (after)")

plot(lc8[[8]], main="June 2014 (before)")
plot(out[[5]], main="June 2014 (after)")
```